
```{r echo=false}
mu.nf=
  c(1.36,1.84,.76,2.25,
    1.41,2.13,1.07,3.64,
    1.77,2.84,1.31,3.13,
    1.02,4.25,1.2,2.88,
    1.21,3.23,1.1,3.96,
    1.22,3.58,.63,3.12,
    1.04,2.62,1.21,2.18,
    1.2,1.41,.79,3.39,
    1.07,1.33,1.55,2.87,
    .68,1.31,.03,-.01,
    .65,1.68,.84,3.55,
    1.21,1.74,.95,5.51,
    1.69,2.61,.74,14.63,
    2.08,3.66,3.06,22.78,
    4.02,2.44,1.93,11.55,
    2.13,2.33,2.76,7.92,
    1.17,4.5,2.09,7.33,
    1.32,1.15,1.01,2.71,
    1.11,.67,1.03,5.79,
    2.4,1.48,1.97,8.7,
    3.78,2.34,1.91)
mu.nf=log(1+(mu.nf/100)) ## convert to continuously compounded returns
nfrmat=matrix(c(mu.nf,NA),(1+length(mu.nf))/4,4,byrow=TRUE)
colnames(nfrmat)=c("Q1","Q2","Q3","Q4")
rownames(nfrmat)=1992:2012


library("zoo")  #load function for reading csv file
ag.df=read.csv(file="Ag commodities.csv")  ## this is dependent on local environment
agidx=ag.df[,2] #extract prices from the data frame
mu.agidx=log(agidx[-1]/agidx[-length(agidx)]) ## calculate continuously compounded monthly returns
mu.agidx=mu.agidx[51:(length(mu.agidx)-1)] ## drop  Nov 1987- Dec 1991 data and Oct 2012 to conform time frame
mu.agidx=matrix(mu.agidx,(length(mu.agidx)/3),3,byrow=TRUE) ## step 1 to convert quarters 
mu.agidx=apply(mu.agidx,1,sum) ## convert to quarterly returns
## next food price index
food.df=read.csv(file="Index mundi food price.csv")
foodidx=food.df[,2] #extract prices from the data frame
mu.foodidx=diff(log(foodidx)) ## calculate CC returns
mu.foodidx=mu.foodidx[12:(length(mu.foodidx)-1)] ## drop 1991 and oct 2012 data
mu.foodidx=matrix(mu.foodidx,(length(mu.foodidx)/3),3,byrow=TRUE)
mu.foodidx=apply(mu.foodidx,1,sum)  ## convert to quarterly returns
## finally grain index
grain.df=read.csv(file="DJUBS grains.csv")
grainidx=grain.df[,2]
mu.grainidx=diff(log(grainidx))
mu.grainidx=mu.grainidx[4:length(mu.grainidx)] ## drop 1991 data, these are already quarterly
## CPI
## CPI X Food and Energy
## Food PPI
## Energy PPI
#create a return matrix
rmat=cbind(mu.nf,mu.agidx,mu.foodidx,mu.grainidx)
colnames(rmat)=c("Farmland","RawMaterials","Food","Grain")
rmat=zooreg(rmat,start=c(1992,1),end=c(2012,3),frequency=4)
```
plot the raw returns
```{r}
plot(rmat)
```
compare the 4 indices in pair wise plots
```{r}
pairs(rmat)
```
Doesn't look like Farmland is connected with commodities
Let's do one regression to check stats
calculate a linear regression ag raw materials as x axis and farm prices as y
```{r}
reg.agidx.nf=lm(mu.nf~mu.agidx)
#display summary of regression results
summary(reg.agidx.nf)
```
The summary data support what are the scatter graph suggests.  Little to no connection.  Note low Rsquared and negative coefficient.

plot data with regression line
```{r}
plot(mu.agidx,mu.nf,main="Regression of raw materials and farmland")
intercept=reg.agidx.nf$coef[1]
coeff=reg.agidx.nf$coeff[2]
lower=min(mu.agidx)
upper=max(mu.agidx)
x=seq(lower,upper,.01)
lines(x,intercept+coeff*x,col="red",type="l")
```
Now lets look at farmland and ag commodities as a portfolio
read in functions for portfolio analysis -- the file address is local environment dependant
```{r}
source('~/R/zivot/portfolio_noshorts.r.txt',encoding='UTF-8')
```
plot a growth of dollar chart
```{r}
plot(exp(cumsum(rmat[,"Farmland"])),xlab="Quarters Q1 1992 thru Q3 2012",ylab="Value",
     main="Growth of Dollar",type='l',col='blue',lwd=2,ylim=c(0,10))
lines(exp(cumsum(rmat[,"RawMaterials"])),type='l',col='red',lwd=2)
lines(exp(cumsum(rmat[,"Food"])),type='l',col='green',lwd=2)
lines(exp(cumsum(rmat[,"Grain"])),type='l',col='black',lwd=2)
legend(x='topleft',legend=colnames(rmat),col=c('blue','red','green','black'),lwd=3)
```
Set up model parameters for portfolio analysis
We are going to calculate a tangency portfolio consisting only of the four agricultural assets.  A tangency portfolio is an efficient portfolio that has the highest sharpe ratio.
```{r}
mu.vals=apply(rmat,2,mean)
sigma2.vals=apply(rmat,2,var)
sigma.vals=apply(rmat,2,sd)
cov.mat=var(rmat)
cor.mat=cor(rmat)
#calculate the tangency portfolio (max sharpe ratio) of ag land and ag commodities
#allow short sales
rf=.005 # risk free rate = .5% per quarter
tan.port=tangency.portfolio(mu.vals,cov.mat,rf,shorts=TRUE)
summary(tan.port)
plot(tan.port, col="green")
```
Note this portfolio has a short position in grains.  Now let's do it prohibiting shorts
```{r}
tan.port.ns=tangency.portfolio(mu.vals,cov.mat,rf,shorts=FALSE)
summary(tan.port.ns)
plot(tan.port.ns, col="green")
```
Let's now consider a portfolio of farmland, stocks and bonds
Note: I'm doing this while traveling. I don't readily have good index data on home computer.  So. . .
Using SPY S&P 500 etf for stocks
Using VFICX Vanguard intermediate term investment grade mutual fund for bonds
In this case downloading prices from Yahoo finance
```{r}
library(tseries)
## download prices from Yahoo
VFICX.prices = get.hist.quote(instrument="vficx", start="1996-07-01",
                              end="2012-10-31", quote="AdjClose",
                              provider="yahoo", origin="1970-01-01",
                              compression="m", retclass="zoo")
SPY.prices = get.hist.quote(instrument="spy", start="1996-07-01",
                              end="2012-10-31", quote="AdjClose",
                              provider="yahoo", origin="1970-01-01",
                              compression="m", retclass="zoo")
## Calculate CC returns
mu.VFICX=diff(log(VFICX.prices))
mu.SPY=diff(log(SPY.prices))
## convert to quarterly
nquarters=length(mu.VFICX)/3
mu.VFICX=matrix(mu.VFICX,nquarters,3,byrow=TRUE)
mu.VFICX=apply(mu.VFICX,1,sum)
mu.SPY=matrix(mu.SPY,nquarters,3,byrow=TRUE)
mu.SPY=apply(mu.SPY,1,sum)
## delete early quarters on NCREIF farm index to match data available on other indices
nquarter.nf=length(mu.nf)
mu.nf2=mu.nf[(1+nquarter.nf-nquarters):nquarter.nf]
rmat=cbind(mu.nf2,mu.SPY,mu.VFICX)
colnames(rmat)=c("Farmland","Stocks","Bonds")
rmat=zooreg(rmat,start=c(1996,3),end=c(2012,3),frequency=4)
```
plot the raw returns
```{r}
plot(rmat)
```
plot the pairwise scattercharts
```{r}
pairs(rmat)
```
plot the returns as growth of dollar chart
```{r}
plot(exp(cumsum(rmat[,"Farmland"])),xlab="Quarters since 1996 Q3",ylab="Value",
                main="Growth of Dollar",type='l',col='blue',lwd=2)
lines(exp(cumsum(rmat[,"Stocks"])),type='l',col='red',lwd=2)
lines(exp(cumsum(rmat[,"Bonds"])),type='l',col='green',lwd=2)
legend(x='topleft',legend=colnames(rmat),col=c('blue','red','green'),lwd=3)
```
Now calculate the tangency portfolio with short sales permitted
```{r}
mu.vals=apply(rmat,2,mean)
mu.vals
sigma2.vals=apply(rmat,2,var)
sigma.vals=apply(rmat,2,sd)
sigma.vals
cov.mat=var(rmat)
cor.mat=cor(rmat)
#allow short sales
rf=.005 # risk free rate = .5% per quarter
tan.port=tangency.portfolio(mu.vals,cov.mat,rf,shorts=TRUE)
summary(tan.port)
```
```{r}
plot(tan.port, col="green")
```
Same thing, but this time prohibiting shorts
```{r}
tan.port.ns=tangency.portfolio(mu.vals,cov.mat,rf,shorts=FALSE)
summary(tan.port.ns)
```
```{r}
plot(tan.port.ns, col="green")
```
This is not a conventional portfolio.
Now let's consider how stable these results are over time.
So, let's do 5 year rolling calculations of tangency portfolio with Shorts allowed
```{r}
rolltan = function(x) {
  mu.hat = colMeans(x)
  cov.hat = var(x)
  gmin = tangency.portfolio(er=mu.hat,cov.mat=cov.hat,risk.free=rf,shorts=TRUE)
  ans = c(gmin$er,gmin$sd,gmin$weights)
  names(ans)[1:2] = c("er","sd")
  ans
}
roll.tan = rollapply(rmat,width=20,
                      by.column=F,align="right",
                      FUN=rolltan)

plot(roll.tan[,3:5],main="Rolling weights in tangency portfolio",
     plot.type="single", col=1:3, lwd=3, ylab="weight")    
abline(h=0)
legend(x="bottomleft",legend=colnames(rmat),
       lty=rep(1,3),col=1:3,lwd=3)
```
```{r}
plot(roll.tan[,1:2],plot.type="single",ylab="percent",
     main="Rolling means and sds on tangency portfolio",
     col=c("black","orange"),lwd=3)
abline(h=0)
legend(x="topleft",legend=colnames(roll.tan[,1:2]),
       lty=rep(1,2),col=c("black","orange"),lwd=3)
```
Now, let's do rolling calculations of tangency portfolio with Shorts prohibited
```{r}
rolltan = function(x) {
  mu.hat = colMeans(x)
  cov.hat = var(x)
  gmin = tangency.portfolio(er=mu.hat,cov.mat=cov.hat,risk.free=rf,shorts=FALSE)
  ans = c(gmin$er,gmin$sd,gmin$weights)
  names(ans)[1:2] = c("er","sd")
  ans
}
roll.tan = rollapply(rmat,width=20,
                     by.column=F,align="right",
                     FUN=rolltan)

plot(roll.tan[,3:5],main="Rolling weights in tangency portfolio",
     plot.type="single", col=1:3, lwd=3, ylab="weight")    
abline(h=0)
legend(x="bottomleft",legend=colnames(rmat),
       lty=rep(1,3),col=1:3,lwd=3)
```
```{r}
plot(roll.tan[,1:2],plot.type="single",ylab="percent",
     main="Rolling means and sds on tangency portfolio",
     col=c("black","orange"),lwd=3)
abline(h=0)
legend(x="topleft",legend=colnames(roll.tan[,1:2]),
       lty=rep(1,2),col=c("black","orange"),lwd=3)
```




