Statistical exploration of the NCREIF Farmland Index
========================================================
ncreif publishes an index of returns of farmland in institutional ownership.  
in this text, we will exam some of the properties of it
Note: all returns throughout are quarterly, continuously compounded

In the text below, the shaded areas are the computer program that generates the graphs and tables presented below.

First, load in the returns.
ncreif farmland returns as %, 1992Q1 thru 2012Q3
variable mu.nf refers to ncreif farmland returns
load in the data and display it (not downloadable, so typed)
```{r}
mu.nf=
  c(1.36,1.84,.76,2.25,
    1.41,2.13,1.07,3.64,
    1.77,2.84,1.31,3.13,
    1.02,4.25,1.2,2.88,
    1.21,3.23,1.1,3.96,
    1.22,3.58,.63,3.12,
    1.04,2.62,1.21,2.18,
    1.2,1.41,.79,3.39,
    1.07,1.33,1.55,2.87,
    .68,1.31,.03,-.01,
    .65,1.68,.84,3.55,
    1.21,1.74,.95,5.51,
    1.69,2.61,.74,14.63,
    2.08,3.66,3.06,22.78,
    4.02,2.44,1.93,11.55,
    2.13,2.33,2.76,7.92,
    1.17,4.5,2.09,7.33,
    1.32,1.15,1.01,2.71,
    1.11,.67,1.03,5.79,
    2.4,1.48,1.97,8.7,
    3.78,2.34,1.91)
mu.nf=log(1+(mu.nf/100)) ## convert to continuously compounded returns
nfrmat=matrix(c(mu.nf,NA),(1+length(mu.nf))/4,4,byrow=TRUE)
colnames(nfrmat)=c("Q1","Q2","Q3","Q4")
rownames(nfrmat)=1992:2012
nfrmat
```
Show some summary statistics and a histogram
```{r}
summary.nf=as.matrix(c(mean(mu.nf),median(mu.nf),sd(mu.nf),max(mu.nf),min(mu.nf)))
rownames(summary.nf)=c("Mean","Median","Standard Deviation","Highest","Lowest")
colnames(summary.nf)="Farmland Quarterly Returns"
summary.nf
hist(mu.nf)
```
```{r}
#create a 9/30 fiscal year matrix drop the 1992 Q1-3 data to get even # of years
#ffy is farmland fiscal years
#nyears is number of years of quarterly data
nyears=(length(mu.nf)-3)/4
ffy=matrix(mu.nf[4:length(mu.nf)],nyears,4,byrow=TRUE)
rownames(ffy)=1993:2012
colnames(ffy)=c("Q4","Q1","Q2","Q3")
```
#calculate the means by quarter
```{r}
qmeans=apply(ffy,2,mean)
names(qmeans)=c("Q4","Q1","Q2","Q3")
qmeans
```
#note the q4 mean is substantially higher than the other quarters
#calculate annual returns (9/30 fiscal years)
```{r}
ffya=as.matrix(apply(ffy,1,sum))
rownames(ffya)=1993:2012
colnames(ffya)="Annual returns"
ffya
```
test is standard deviation consistent with standard deviation of quarterly data
```{r}
ffya_sd=sd(ffya[,1])
nfr_sd=sd(mu.nf)

## expected annual standard deviation from quarterly data is sqrt(4) * the quarterly sd
## show the difference in a table
sd_test=as.matrix(c(nfr_sd,2*nfr_sd,ffya_sd))
rownames(sd_test)=c("Calcuated quarterly sd", "Implied annual sd from quarterly",
                    "Calculated annual sd")
colnames(sd_test)="Standard Deviations"
sd_test
```                  
The standard deviation calculated from annual data is not very different from the annual standard deviation implied by the quarterly data, but there is clearly a bias in Q4 returns, presumably because of valuation practices at year end.

now load agricultural commodities indices
in this case they are in an excel file -- loading from the file  
```{r}
library("zoo")  #load function for reading csv file
```
Load various ag indeces
Raw materials index is cotton, hides -- ag products used in making things
Food index
Grains index
```{r}
ag.df=read.csv(file="C:/Users/LR/Dropbox/R/Ag commodities.csv")  ## this is dependent on local environment
agidx=ag.df[,2] #extract prices from the data frame
mu.agidx=log(agidx[-1]/agidx[-length(agidx)]) ## calculate continuously compounded monthly returns
mu.agidx=mu.agidx[51:(length(mu.agidx)-1)] ## drop  Nov 1987- Dec 1991 data and Oct 2012 to conform time frame
mu.agidx=matrix(mu.agidx,(length(mu.agidx)/3),3,byrow=TRUE) ## step 1 to convert quarters 
mu.agidx=apply(mu.agidx,1,sum) ## convert to quarterly returns
## next food price index
food.df=read.csv(file="c:/users/LR/Dropbox/R/Index mundi food price.csv")
foodidx=food.df[,2] #extract prices from the data frame
mu.foodidx=diff(log(foodidx)) ## calculate CC returns
mu.foodidx=mu.foodidx[12:(length(mu.foodidx)-1)] ## drop 1991 and oct 2012 data
mu.foodidx=matrix(mu.foodidx,(length(mu.foodidx)/3),3,byrow=TRUE)
mu.foodidx=apply(mu.foodidx,1,sum)  ## convert to quarterly returns
## finally grain index
grain.df=read.csv(file="c:/users/LR/Dropbox/R/DJUBS grains.csv")
grainidx=grain.df[,2]
mu.grainidx=diff(log(grainidx))
mu.grainidx=mu.grainidx[4:length(mu.grainidx)] ## drop 1991 data, these are already quarterly
#create a return matrix
rmat=cbind(mu.nf,mu.agidx,mu.foodidx,mu.grainidx)
colnames(rmat)=c("Farmland","RawMaterials","Food","Grain")
rmat=zooreg(rmat,start=c(1992,1),end=c(2012,3),frequency=4)
```
plot the raw returns
```{r}
plot(rmat)
```
compare the 4 indices in pair wise plots
```{r}
pairs(rmat)
```
Doesn't look like Farmland is connected with commodities
Let's do one regression to check stats
calculate a linear regression ag raw materials as x axis and farm prices as y
```{r}
reg.agidx.nf=lm(mu.nf~mu.agidx)
#display summary of regression results
summary(reg.agidx.nf)
```
The summary data support what are the scatter graph suggests.  Little to no connection.  Note low Rsquared and negative coefficient.

plot data with regression line
```{r}
plot(mu.agidx,mu.nf,main="Regression of raw materials and farmland")
intercept=reg.agidx.nf$coef[1]
coeff=reg.agidx.nf$coeff[2]
lower=min(mu.agidx)
upper=max(mu.agidx)
x=seq(lower,upper,.01)
lines(x,intercept+coeff*x,col="red",type="l")
```
Now lets look at farmland and ag commodities as a portfolio
read in functions for portfolio analysis -- the file address is local environment dependant
```{r}
source('~/R/zivot/portfolio_noshorts.r.txt',encoding='UTF-8')
```
plot a growth of dollar chart
```{r}
plot(exp(cumsum(rmat[,"Farmland"])),xlab="Quarters Q1 1992 thru Q3 2012",ylab="Value",
     main="Growth of Dollar",type='l',col='blue',lwd=2,ylim=c(0,10))
lines(exp(cumsum(rmat[,"RawMaterials"])),type='l',col='red',lwd=2)
lines(exp(cumsum(rmat[,"Food"])),type='l',col='green',lwd=2)
lines(exp(cumsum(rmat[,"Grain"])),type='l',col='black',lwd=2)
legend(x='topleft',legend=colnames(rmat),col=c('blue','red','green','black'),lwd=3)
```
Set up model parameters for portfolio analysis
We are going to calculate a tangency portfolio consisting only of the four agricultural assets.  A tangency portfolio is an efficient portfolio that has the highest sharpe ratio.
```{r}
mu.vals=apply(rmat,2,mean)
sigma2.vals=apply(rmat,2,var)
sigma.vals=apply(rmat,2,sd)
cov.mat=var(rmat)
cor.mat=cor(rmat)
#calculate the tangency portfolio (max sharpe ratio) of ag land and ag commodities
#allow short sales
rf=.005 # risk free rate = .5% per quarter
tan.port=tangency.portfolio(mu.vals,cov.mat,rf,shorts=TRUE)
summary(tan.port)
plot(tan.port, col="green")
```
Note this portfolio has a short position in grains.  Now let's do it prohibiting shorts
```{r}
tan.port.ns=tangency.portfolio(mu.vals,cov.mat,rf,shorts=FALSE)
summary(tan.port.ns)
plot(tan.port.ns, col="green")
```
Let's now consider a portfolio of farmland, stocks and bonds
Note: I'm doing this while traveling. I don't readily have good index data on home computer.  So. . .
Using SPY S&P 500 etf for stocks
Using VFICX Vanguard intermediate term investment grade mutual fund for bonds
In this case downloading prices from Yahoo finance
```{r}
library(tseries)
## download prices from Yahoo
VFICX.prices = get.hist.quote(instrument="vficx", start="1996-07-01",
                              end="2012-10-31", quote="AdjClose",
                              provider="yahoo", origin="1970-01-01",
                              compression="m", retclass="zoo")
SPY.prices = get.hist.quote(instrument="spy", start="1996-07-01",
                              end="2012-10-31", quote="AdjClose",
                              provider="yahoo", origin="1970-01-01",
                              compression="m", retclass="zoo")
## Calculate CC returns
mu.VFICX=diff(log(VFICX.prices))
mu.SPY=diff(log(SPY.prices))
## convert to quarterly
nquarters=length(mu.VFICX)/3
mu.VFICX=matrix(mu.VFICX,nquarters,3,byrow=TRUE)
mu.VFICX=apply(mu.VFICX,1,sum)
mu.SPY=matrix(mu.SPY,nquarters,3,byrow=TRUE)
mu.SPY=apply(mu.SPY,1,sum)
## delete early quarters on NCREIF farm index to match data available on other indices
nquarter.nf=length(mu.nf)
mu.nf2=mu.nf[(1+nquarter.nf-nquarters):nquarter.nf]
rmat=cbind(mu.nf2,mu.SPY,mu.VFICX)
colnames(rmat)=c("Farmland","Stocks","Bonds")
rmat=zooreg(rmat,start=c(1996,3),end=c(2012,3),frequency=4)
```
plot the raw returns
```{r}
plot(rmat)
```
plot the pairwise scattercharts
```{r}
pairs(rmat)
```
plot the returns as growth of dollar chart
```{r}
plot(exp(cumsum(rmat[,"Farmland"])),xlab="Quarters since 1996 Q3",ylab="Value",
                main="Growth of Dollar",type='l',col='blue',lwd=2)
lines(exp(cumsum(rmat[,"Stocks"])),type='l',col='red',lwd=2)
lines(exp(cumsum(rmat[,"Bonds"])),type='l',col='green',lwd=2)
legend(x='topleft',legend=colnames(rmat),col=c('blue','red','green'),lwd=3)
```
Now calculate the tangency portfolio with short sales permitted
```{r}
mu.vals=apply(rmat,2,mean)
mu.vals
sigma2.vals=apply(rmat,2,var)
sigma.vals=apply(rmat,2,sd)
sigma.vals
cov.mat=var(rmat)
cor.mat=cor(rmat)
#allow short sales
rf=.005 # risk free rate = .5% per quarter
tan.port=tangency.portfolio(mu.vals,cov.mat,rf,shorts=TRUE)
summary(tan.port)
```
```{r}
plot(tan.port, col="green")
```
Same thing, but this time prohibiting shorts
```{r}
tan.port.ns=tangency.portfolio(mu.vals,cov.mat,rf,shorts=FALSE)
summary(tan.port.ns)
```
```{r}
plot(tan.port.ns, col="green")
```
This is not a conventional portfolio.
Now let's consider how stable these results are over time.
So, let's do 5 year rolling calculations of tangency portfolio with Shorts allowed
```{r}
rolltan = function(x) {
  mu.hat = colMeans(x)
  cov.hat = var(x)
  gmin = tangency.portfolio(er=mu.hat,cov.mat=cov.hat,risk.free=rf,shorts=TRUE)
  ans = c(gmin$er,gmin$sd,gmin$weights)
  names(ans)[1:2] = c("er","sd")
  ans
}
roll.tan = rollapply(rmat,width=20,
                      by.column=F,align="right",
                      FUN=rolltan)

plot(roll.tan[,3:5],main="Rolling weights in tangency portfolio",
     plot.type="single", col=1:3, lwd=3, ylab="weight")    
abline(h=0)
legend(x="bottomleft",legend=colnames(rmat),
       lty=rep(1,3),col=1:3,lwd=3)
```
```{r}
plot(roll.tan[,1:2],plot.type="single",ylab="percent",
     main="Rolling means and sds on tangency portfolio",
     col=c("black","orange"),lwd=3)
abline(h=0)
legend(x="topleft",legend=colnames(roll.tan[,1:2]),
       lty=rep(1,2),col=c("black","orange"),lwd=3)
```
Now, let's do rolling calculations of tangency portfolio with Shorts prohibited
```{r}
rolltan = function(x) {
  mu.hat = colMeans(x)
  cov.hat = var(x)
  gmin = tangency.portfolio(er=mu.hat,cov.mat=cov.hat,risk.free=rf,shorts=FALSE)
  ans = c(gmin$er,gmin$sd,gmin$weights)
  names(ans)[1:2] = c("er","sd")
  ans
}
roll.tan = rollapply(rmat,width=20,
                     by.column=F,align="right",
                     FUN=rolltan)

plot(roll.tan[,3:5],main="Rolling weights in tangency portfolio",
     plot.type="single", col=1:3, lwd=3, ylab="weight")    
abline(h=0)
legend(x="bottomleft",legend=colnames(rmat),
       lty=rep(1,3),col=1:3,lwd=3)
```
```{r}
plot(roll.tan[,1:2],plot.type="single",ylab="percent",
     main="Rolling means and sds on tangency portfolio",
     col=c("black","orange"),lwd=3)
abline(h=0)
legend(x="topleft",legend=colnames(roll.tan[,1:2]),
       lty=rep(1,2),col=c("black","orange"),lwd=3)
```




